---
title: "GLM for 5523"
author: "Lukas Buhler"
date: "12/1/2021"
output: html_document
---

Loading Data
```{r}
Data<-read.csv("C:\\Users\\buhl5\\Documents\\StatData.csv")
str(Data)
```
We have many values read in as characters that need to be converted to factors
for modeling

Factorization
```{r}
names(Data)
Data$contact<- factor(Data$contact)
Data$Minneapolis<- factor(Data$Minneapolis)
Data$RacialGroup<- factor(Data$RacialGroup, levels = c("White American", "Somali American", "African American"))
Data$Type<- factor(Data$Type)
Data$Sex<- factor(Data$Sex)
Data$Law<- factor(Data$Law)
Data$batch <- factor(Data$batch)

levels(Data$contact)<-c("Ambiguous","Ghosted","Negative","Positive")
```

Exploring our variables
```{r}
with(Data,plot(factor(Data$Minneapolis),factor(Data$contact)))
with(Data,plot(factor(Data$RacialGroup),factor(Data$contact)))
with(Data,plot(factor(Data$Law),factor(Data$contact)))
with(Data,plot(factor(Data$Sex),factor(Data$contact)))
with(Data,plot(factor(Data$StayAtHome),factor(Data$contact)))
with(Data,plot(factor(Data$Type),factor(Data$contact)))
plot(Data$contact)
summary(Data$contact)
```

Variable Selection
```{r}
nmod<- glm(Positive~1,family=binomial,data=Data)
memod <-glm(Positive~Type+MonthlyRent+Size+RacialGroup+Sex+Minneapolis+Law+StayAtHome,family=binomial,data=Data)
```

Forward Selection
```{r}
library(MASS)
stepAIC(nmod, scope = list(upper=~Type+MonthlyRent+Size+RacialGroup+Sex+Minneapolis+Law+StayAtHome, lower=~1),direction="both",trace=0,k=2,data=Data)
```

Backward Selection
```{r}
stepAIC(memod, scope = list(upper=~Type+MonthlyRent+Size+RacialGroup+Sex+Minneapolis+Law+StayAtHome, lower=~1),direction="both",trace=0,k=2,data=Data)
```
Both directions agreed that Postive~Type+MonthlyRent+Size+RacialGroup+Sex+Law is
the best main effects model. We may want to try interactions though. As we know
the law variable is directly related to the racial identity of applicants, so
we will try an interaction term.

```{r}
mod1 <- glm(Positive~Type+MonthlyRent+Size+RacialGroup*Law+Sex,family=binomial,data=Data)
summary(mod1)
```
The AIC did not improve with the new model so we will not replace the old model.

Sex and RacialGroup interaction
```{r}
mod2 <- glm(Positive~Type+MonthlyRent+Size+RacialGroup*Sex+Law,family=binomial,data=Data)
summary(mod2)
```
The interaction of Sex and Race does not improve the AIC score either. We will
continue with the model chosen by stepAIC as there are no other interactions we
could consider meaningful in an applied context.

Prediction Threshold
```{r}
mod <- glm(Positive~Type+MonthlyRent+Size+RacialGroup+Sex+Law,family=binomial,data=Data)
```

Testing prediction based on accuracy using different thresholds
```{r}
out <- predict(mod,type="response") 
summary(out) #our model is conservative in prediction never giving more than .6 for p
threshold <- seq(0.2,0.61,0.01) #just below lowest to just above highest
prop <- rep(0,length(threshold))
c <- rep(0,length(threshold))
recall <- rep(0,length(threshold))
total <- nrow(Data)
for (k in 1:length(threshold)) {
  pred <- ifelse(out > threshold[k],1,0)
  pred2 <- ifelse(out>threshold[k],3,-3)
  tab <- table(Data$Positive-pred2)
  val <- ifelse(pred == Data$Positive,1,0)
  c[k] <- sum(abs(val))
  prop[k] <- c[k]/total
  recall[k] <- tab[4]/(tab[4]+tab[2])
}
prop
recall
```

```{r}
par(mfrow=c(1,2))
plot(prop)
plot(recall)
max(recall,na.rm=TRUE)
prop[37]
recall[37]
```
The final model will use a prediction threshold of 0.56 as it produced the highest
recall without classifying all values as positive. It also has an overall accuracy of
0.6423 which is not great. However I feel that it is more important to get the positives
and over predict to encourage people to apply anyways as opposed to potentially
missing positives and having people not apply for housing they may have gotten a 
lease on.




